---
title: "nDrichr Report"
author: "Antony Kaspi & Mark Ziemann"
runtime: shiny
output: html_document
theme: cosmo
---

date generated: "`r Sys.Date()`"

## Background

nDrich performs multidimensional gene set enrichment analysis. The concept behind this dates to work by Cox and 
Mann (https://doi.org/10.1186/1471-2105-13-S16-S12). This implementation is suited to R based workflows of 
multi-omics datasets. This software was developed by Antony Kaspi and Mark Ziemann.

## Input profiles

Here is the first few lines of the input profile.

```{r,peek,echo=FALSE}
load(DATANAME)
head(res$input_profile)
```

Here are some metrics about the input data profile:

```{r, metrics, echo=FALSE}
unformatted<-t(as.data.frame(res$manova_analysis_metrics[ c(2,3,4,5,11,12 ) ])) 
formatted<-unformatted
formatted[1:4]<-as.character(round(as.numeric(unformatted[1:4]) , digits=0))
formatted[5:6]<-as.character(round(as.numeric(unformatted[5:6]) , digits=5))
colnames(formatted)="Profile metrics"
kable( formatted, caption = "Profile metrics" )
```

Here is a scatterplot of the input profiles. Note the dynamic ranges.

```{r, scatterplot, echo=FALSE,fig.height = 6, fig.width = 6.5, message=FALSE}
if ( ncol(res$input_profile)<3 ) {
  plot(res$input_profile, pch=19, col=rgb(red = 0, green = 0, blue = 0, alpha = 0.15), main="Input profiles"  )
} else {
  library("GGally")
  library("vioplot")
  ggpairs_points_plot <- function(data ,mapping, ...){
    p <- ggplot(data = data, mapping = mapping) +
    geom_point(alpha=0.05) +
    geom_vline(xintercept=0,linetype="dashed") +
    geom_hline(yintercept=0,linetype="dashed")
  }
  p<-ggpairs(as.data.frame(res$input_profile), title="Scatterplot of all genes" , lower  = list(continuous = ggpairs_points_plot ))
  print( p +  theme_bw() )
}
```

Here is the contour plot of the profile including all genes.

```{r, contourplot, echo=FALSE,fig.height = 6, fig.width = 6.5}
palette <- colorRampPalette(c("white", "yellow","orange" ,"red","darkred","black"))
#Contour of all the data
ss<-res$ranked_profile
xmin=min(ss[,1])
xmax=max(ss[,1])
ymin=min(ss[,2])
ymax=max(ss[,2])
ss<-res$ranked_profile
if ( ncol(ss)<3 ) {
  k<-MASS:::kde2d(ss[,1],ss[,2])
  X_AXIS=paste("Rank in contrast",colnames(ss)[1])
  Y_AXIS=paste("Rank in contrast",colnames(ss)[2])
  filled.contour(k, xlim=c(xmin,xmax),ylim=c(ymin,ymax),
    color=palette ,
    plot.title={ abline(v=0,h=0,lty=2,lwd=2,col="blue")
      title( main="Rank-rank plot of all genes",xlab=X_AXIS,ylab=Y_AXIS ) } )
} else {
  #pairs contour plot function
  ggpairs_func <- function(data, mapping, ...){
    p <- ggplot(data = data, mapping = mapping) +
      stat_density2d(aes(fill=..density..), geom="tile", contour = FALSE) +
      geom_vline(xintercept=0,linetype="dashed") +
      geom_hline(yintercept=0,linetype="dashed") +
      scale_fill_gradientn(colours=palette(25))
    p
  }

  #pairs contour plot
  p<-ggpairs(as.data.frame(ss), title="Contour plot of all genes after ranking" , lower=list(continuous=ggpairs_func),
    diag=list(continuous=wrap("barDiag", binwidth=nrow(ss)/100)))
  print( p + theme_bw() )
  #subset contour plot
  ggpairs_contour_limit_range <- function(data ,mapping, ...){
    p <- ggplot(data = data, mapping = mapping) +
      stat_density2d(aes(fill=..density..), geom="tile", contour = FALSE) +
      geom_vline(xintercept=0,linetype="dashed") +
      geom_hline(yintercept=0,linetype="dashed") +
      scale_fill_gradientn(colours=palette(25)) +
      scale_x_continuous( limits = range(min(ss[,gsub("~","",as.character(mapping[1]))]),max(ss[,gsub("~","",as.character(mapping[1]))])) ) +
      scale_y_continuous( limits = range(min(ss[,gsub("~","",as.character(mapping[2]))]),max(ss[,gsub("~","",as.character(mapping[2]))])) )
    p
  }

  #subset points plot
  ggpairs_points_limit_range <- function(data ,mapping, ...){
    p <- ggplot(data = data, mapping = mapping) +
      geom_point(alpha=0.1) +
      geom_vline(xintercept=0,linetype="dashed") +
      geom_hline(yintercept=0,linetype="dashed") +
      scale_x_continuous( limits = range(min(ss[,gsub("~","",as.character(mapping[1]))]),max(ss[,gsub("~","",as.character(mapping[1]))])) ) +
      scale_y_continuous( limits = range(min(ss[,gsub("~","",as.character(mapping[2]))]),max(ss[,gsub("~","",as.character(mapping[2]))])) )
    p
  }
}
```

Here is the distribution of genes in the quadrants/sectors
```{r, quadrant_distribution, results="asis",echo=FALSE}
if ( ncol(ss)<3 ) {
  uu=length(which(res$input_profile[,1]>0 & res$input_profile[,2]>0))
  ud=length(which(res$input_profile[,1]>0 & res$input_profile[,2]<0))
  dd=length(which(res$input_profile[,1]<0 & res$input_profile[,2]<0))
  du=length(which(res$input_profile[,1]<0 & res$input_profile[,2]>0))
  a<-as.data.frame(c(uu,ud,dd,du))
  rownames(a)=c("top-right","bottom-right","bottom-left","top-left")
  colnames(a)="a"
  barplot(a$a,names.arg=rownames(a),main="number of genes in each quadrant")
} else {
  d=ncol(ss)
  sig<-sign(ss)
  sector_count<-aggregate(1:nrow(sig) ~ ., sig, FUN = length)
  colnames(sector_count)[ncol(sector_count)]<-"Count"
  kable(sector_count ,caption = "Genes by sector" ,row.names=T)
}

```

## Input genesets
Here are some metrics about the gene sets used:
```{r,input_geneset_metrics,results="asis",echo=FALSE}
ORIGINFILE=attributes(genesets)$originfile
print(paste("GMT file of genesets:",ORIGINFILE))
unformatted<-t(as.data.frame(res$manova_analysis_metrics[ c(1,6,7,13  ) ]))
formatted<-as.data.frame(as.character( unformatted[1:4]) )
rownames(formatted)=rownames(unformatted)
colnames(formatted)="Gene sets metrics"
kable( formatted , col.names = "Gene sets metrics" ,caption = "Gene sets metrics" )

if ( ncol(ss)<3 ) {
  a<-res$manova_analysis_metrics[14]
  a<-as.data.frame(as.numeric(unlist(strsplit(as.character(a),','))),stringsAsFactors=F)
  rownames(a)=c("top-right","bottom-right","bottom-left","top-left")
  colnames(a)="a"
  barplot(a$a,names.arg=rownames(a),main="number of genesets FDR<0.05")
} 
d=ncol(ss)
sig<-sign(res$manova_result[which(res$manova_result$p.adjustMANOVA<0.05),4:(4+d-1)])
sector_count<-aggregate(1:nrow(sig) ~ ., sig, FUN = length)
colnames(sector_count)[ncol(sector_count)]<-"Count"
kable(sector_count ,caption = "Gene sets by sector" ,row.names=T)
```

More about the sizes of the gene sets evaluated.
```{r,geneset_sizes,results="asis",echo=FALSE,fig.height = 10, fig.width = 7, fig.show="all"}
par(mfrow=c(3,1))
geneset_counts<-res$manova_analysis_metrics$geneset_counts
boxplot(geneset_counts$count,horizontal=T,frame=F,main="Gene set size",xlab="number of member genes included in profile")
hist(geneset_counts$count,100,xlab="geneset size",main="Histogram of geneset size")
hist(geneset_counts$count,100,xlim=c(0,500),xlab="geneset size",main="Trimmed histogram of geneset size")
```

## Enrichment scatterplot with taucharts
All sets with p.adjustMANOVA<0.05. Try hovering over the points.
```{r,tauchart1,results="asis",echo=F, fig.height = 7, fig.width = 7 ,fig.show="all", message=FALSE}
library("taucharts")
library(dplyr)
library(gtools)
  numsets=nrow(subset(res$manova_result,p.adjustMANOVA<0.05))
  p=NULL
  if (numsets>0){
    myres<-subset(res$manova_result,p.adjustMANOVA<0.05)
    myres$p.adjustMANOVA<-signif(myres$p.adjustMANOVA,3)
    if ( ncol(ss)<3 ) {
      myres<-myres[,c(1,8,2,4,5)]
      myres$set<-gsub("_"," ",as.character(myres$set))
      colnames(myres)=gsub("\\.","_",colnames(myres))
      my_x=colnames(myres)[4]
      my_y=colnames(myres)[5]    
      tauchart(myres) %>% tau_point(my_x,my_y) %>% tau_tooltip()
    } else {
      plan<-combinations(n = d, r = 2, v = 1:d, repeats.allowed = FALSE)
      myres<-myres[,c(1,ncol(myres)-1,2,4:(4+d-1))]
      p<-list()
      tauscatter<-function(i){
        my_x=plan[i,1]
        my_y=plan[i,2]
        colnames(myres)=gsub("\\.","_",colnames(myres))
        my_x=colnames(myres)[3+my_x]
        my_y=colnames(myres)[3+my_y]
        myres$set<-gsub("_"," ",as.character(myres$set))
        tauchart(myres) %>% tau_point(my_x,my_y) %>% tau_tooltip() 
      }
      p<-lapply( 1:nrow(plan) , tauscatter)
    }
  } else {
    message("No significant enrichments found.")
  }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
```

## Enrichment scatterplot with taucharts
Top N sets irrespective of FDR. Try hovering over the points.
```{r,tauchart2,results="asis",echo=F, fig.height = 7, fig.width = 7 ,fig.show="all", message=FALSE}
library("taucharts")
library(dplyr)
library(gtools)
  resrows=length(res$detailed_sets)
  myres<-head(res$manova_result,resrows)
  myres$p.adjustMANOVA<-signif(myres$p.adjustMANOVA,3)
  if ( ncol(ss)<3 ) {
    myres<-myres[,c(1,10,2,4,5)]
    myres$set<-gsub("_"," ",as.character(myres$set))
    colnames(myres)=gsub("\\.","_",colnames(myres))
    my_x=colnames(myres)[4]
    my_y=colnames(myres)[5]
    tauchart(myres) %>% tau_point(my_x,my_y) %>% tau_tooltip()
  } else {
    plan<-combinations(n = d, r = 2, v = 1:d, repeats.allowed = FALSE)
    myres<-myres[,c(1,ncol(myres),2,4:(4+d-1))]
    p<-list()
    tauscatter<-function(i){
      my_x=plan[i,1]
      my_y=plan[i,2]
      colnames(myres)=gsub("\\.","_",colnames(myres))
      my_x=colnames(myres)[3+my_x]
      my_y=colnames(myres)[3+my_y]
      myres$set<-gsub("_"," ",as.character(myres$set))
      tauchart(myres) %>% tau_point(my_x,my_y) %>% tau_tooltip()
    }
    p<-lapply( 1:nrow(plan) , tauscatter)
  }

if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
```

# A plot of effect size versus significance
Significance is the -log2(p.adjustMANOVA) and effect size is the s.dist which is the hypotenuse of the s scores.
```{r,effectsize,results="asis",echo=F, fig.height = 7, fig.width = 7 ,fig.show="all", message=FALSE}
  myres<-res$manova_result
  myres$significance<--log2(myres$p.adjustMANOVA)
  myres$set<-gsub("_"," ",as.character(myres$set))
  colnames(myres)=gsub("\\.","_",colnames(myres))
  my_x=colnames(myres)[ncol(myres)-2]
  my_y=colnames(myres)[ncol(myres)]
  tauchart(myres) %>% tau_point(my_x,my_y) %>% tau_tooltip()

```



## Results table
```{r,results_table,results="asis",echo=FALSE}
resrows=length(res$detailed_sets)
myres<-head(res$manova_result,resrows)
myres<-myres[,c(1:3,ncol(myres)-1,4:((d*2)+3))]
formatted<-myres
formatted[,1]<-gsub("_"," ",myres[,1])
formatted[,3:ncol(myres)]<-format(myres[,3:ncol(myres)],digits=3)
kable( formatted , col.names = colnames(formatted) , row.names=F, caption = "Top gene sets" ,digits=100)
cat("<hr><br>")
```

## Results (complete table)
```{r,results_table_complete,results="asis",echo=FALSE}
myres<-res$manova_result
myres<-myres[,c(1:3,ncol(myres)-1,4:((d*2)+3))]
formatted<-myres
formatted[,1]<-gsub("_"," ",myres[,1])
formatted[,3:ncol(myres)]<-format(myres[,3:ncol(myres)],digits=3)
HEADER=paste("<br><details><summary><b>","Click HERE to show results for all gene sets","</b></summary><br><p>",sep=" " )
cat(HEADER)
kable(formatted, col.names=colnames(formatted),row.names=F,caption = "Complete results",digits=100)
cat("<br></p></details>")
cat("<hr><br>")
```

## Detailed Gene set reports

```{r,detailed_geneset_reports1,results="asis",echo=FALSE,fig.height=5, fig.width=6,out.width = '40%',comment=NA, message=FALSE}
library("GGally")
library(vioplot)
resrows=length(res$detailed_sets)
ss<-res$ranked_profile
myres<-head(res$manova_result,resrows)
myres<-myres[,c(1:3,ncol(myres)-1,4:((d*2)+3))]
palette <- colorRampPalette(c("white", "yellow","orange" ,"red","darkred","black"))

twodimplot<-function(i) {
  ll<-res$manova_result[i,]
  size<-ll$setSize
  sss<-res$detailed_sets[[i]]
  k<-MASS:::kde2d(sss[,1],sss[,2])
    filled.contour( k, color = palette, xlim=c(xmin,xmax),ylim=c(ymin,ymax),
    plot.title={ abline(v=0,h=0,lty=2,lwd=2,col="blue")
      title( main=ll$set , xlab=X_AXIS,ylab=Y_AXIS  )
    }
  )
  plot(sss, pch=19, col=rgb(red = 0, green = 0, blue = 0, alpha = 0.2),
    main=ll$set ,
    xlim=c(xmin,xmax),ylim=c(ymin,ymax),
    xlab=X_AXIS,ylab=Y_AXIS
  )
  abline(v=0,h=0,lty=2,lwd=2,col="blue")
  do.call(vioplot,c(unname(as.data.frame(sss)),col='gray',drawRect=T,names=list(names(as.data.frame(sss)))))
  grid()
  abline(h=0,lty=2,lwd=2,col="blue")
  title(main = ll[,1] , ylab = "Position in rank")
}

ndplot<-function(i) {
  ll<-res$manova_result[i,]
  size<-ll$setSize
  sss<-res$detailed_sets[[i]]

  #subset contour plot
  ggpairs_contour_limit_range <- function(data ,mapping, ...){
    p <- ggplot(data = data, mapping = mapping) +
      stat_density2d(aes(fill=..density..), geom="tile", contour = FALSE) +
      geom_vline(xintercept=0,linetype="dashed") +
      geom_hline(yintercept=0,linetype="dashed") +
      scale_fill_gradientn(colours=palette(25)) +
      scale_x_continuous( limits = range(min(ss[,gsub("~","",as.character(mapping[1]))]),max(ss[,gsub("~","",as.character(mapping[1]))])) ) +
      scale_y_continuous( limits = range(min(ss[,gsub("~","",as.character(mapping[2]))]),max(ss[,gsub("~","",as.character(mapping[2]))])) )
    p
  }

  #subset points plot
  ggpairs_points_limit_range <- function(data ,mapping, ...){
    p <- ggplot(data = data, mapping = mapping) +
      geom_point(alpha=0.1) +
      geom_vline(xintercept=0,linetype="dashed") +
      geom_hline(yintercept=0,linetype="dashed") +
      scale_x_continuous( limits = range(min(ss[,gsub("~","",as.character(mapping[1]))]),max(ss[,gsub("~","",as.character(mapping[1]))])) ) +
      scale_y_continuous( limits = range(min(ss[,gsub("~","",as.character(mapping[2]))]),max(ss[,gsub("~","",as.character(mapping[2]))])) )
    p
  }

  p<-ggpairs(as.data.frame(sss), title=ll[,1], lower=list(continuous=ggpairs_contour_limit_range),
    diag=list(continuous=wrap("barDiag", binwidth=nrow(ss)/10)) )
  print( p + theme_bw() )
  p<-ggpairs(as.data.frame(sss), title=ll[,1], lower= list(continuous = ggpairs_points_limit_range ),
    diag=list(continuous=wrap("barDiag", binwidth=nrow(ss)/10)))
  print( p + theme_bw() )
  do.call(vioplot,c(unname(as.data.frame(sss)),col='gray',drawRect=T,names=list(names(as.data.frame(sss)))))
  grid()
  abline(h=0,lty=2,lwd=2)
  title(main = ll[,1] , ylab = "Position in rank")
}

topgenelist<-function(i) {
  sss<-res$detailed_sets[[i]]
  tl=bl=tr=br=0

  d=ncol(ss)
  #select 2 strongest columns to highlight genes
  if( d>2 ) {
    cols<-order(-abs(myres[i,5:(d+4)]))[1:2]
    sss<-sss[,cols]
    cols=cols+4
    mysx=signif(myres[i,cols[1]],3)
    mysy=signif(myres[i,cols[2]],3)
  } else {
    cols=c(5,6)
    mysx=signif(myres[i,cols[1]],3)
    mysy=signif(myres[i,cols[2]],3)
  }

  if ( mysy>0 ) { tl=tl+1 ; tr=tr+1  } else { bl=bl+1 ; br=br+1 }
  if ( mysx>0 ) { tr=tr+1 ; br=br+1  } else { tl=tl+1 ; bl=bl+1 }
  if (bl==2) {
    myquad<-sss[which(sss[,1]<0 & sss[,2]<0),]
    topgenes<-myquad[order(-rank(myquad[,1]*myquad[,2])),]
  }
  if (tr==2) {
    myquad<-sss[which(sss[,1]>0 & sss[,2]>0),]
    topgenes<-myquad[order(-rank(myquad[,1]*myquad[,2])),]
  }
  if (br==2) {
    myquad<-sss[which(sss[,1]>0 & sss[,2]<0),]
    topgenes<-myquad[order(rank(myquad[,1]*myquad[,2])),]
  }
  if (tl==2) {
    myquad<-sss[which(sss[,1]<0 & sss[,2]>0),]
    topgenes<-myquad[order(rank(myquad[,1]*myquad[,2])),]
  }
  cat("<br>")
  topgenes<-as.data.frame(topgenes)
  topgenes$Gene<-as.character(rownames(topgenes))
  topgenes<-topgenes[,c(3,1,2)]
  print(kable(head(topgenes,n=20L),col.names=colnames(topgenes),format="markdown",row.names=F,caption="Top 20 genes",digits=100))
  cat('\n')
  HEADER=paste("<br><details><summary><b>","Click HERE to show all gene set members","</b></summary><br><p>",sep=" " )
  cat(HEADER)
  sss<-res$detailed_sets[[i]]
  print(kable(sss,format="markdown",row.names=T,caption="All member genes",digits=100))
  cat('\n')
  cat("<br></p></details>")
  cat("<br><hr>")
}

#functions end here
if ( ncol(ss)<3 ) {
  xmin=min(ss[,1])
  xmax=max(ss[,1])
  ymin=min(ss[,2])
  ymax=max(ss[,2])
  for ( i in 1:resrows ) {
    mydat<-as.data.frame(t(myres[i,2:ncol(myres)]))
    mydat$' '<-" "
#    mydat$rowname<-rownames(mydat)
#    mydat<-mydat[,c(2,1)]
    cat(paste("<b>",as.character(myres[i,1]) ,"</b><br>"))
    print(kable(mydat,format='markdown',digits=5,caption=as.character(t(myres[i,1]))) )
    cat('\n')
    cat("<br>")
    twodimplot(i)
    cat("<br>")
    topgenelist(i)
    cat("<hr><br>")
  }
} else {
  for ( i in 1:resrows ) {
    mydat<-as.data.frame(t(myres[i,2:ncol(myres)]))
    mydat$' '<-" "
#    mydat$rowname<-rownames(mydat)
#    mydat<-mydat[,c(2,1)]
    cat(paste("<b>",as.character(myres[i,1]) ,"</b><br>"))
    print(kable(mydat,format='markdown',digits=5,caption=as.character(t(myres[i,1]))))
    cat('\n')
    cat("<br>")
    ndplot(i)
    cat("<br>")
    topgenelist(i)
    cat("<hr><br>")
  }
}
```

END of report
