---
title: "Mitch Report"
author: "Antony Kaspi & Mark Ziemann"
output: html_document
theme: cosmo
---

date generated: "`r Sys.Date()`"

## Background

Mitch performs unidimensional and multidimensional gene set enrichment analysis. The
concept behind this dates to work by Cox and Mann 
(https://doi.org/10.1186/1471-2105-13-S16-S12). This implementation is suited to R based 
workflows of multi-omics datasets. This software was developed by Antony Kaspi and Mark 
Ziemann. Learn more about Mitch at the website: https://github.com/markziemann/Mitch

## Input profiles

Here is the first few lines of the input profile.

```{r,peek,echo=FALSE}
load(DATANAME)
  # capture the dimensionality of the data                                               
  d=ncol(res$input_profile)
  if (d==1) { D1=TRUE } else { D1=FALSE }
  if (d==2) { D2=TRUE } else { D2=FALSE }
  if (d>2 & d<=5) { D3=TRUE } else {D3=FALSE }
  if (d>5) { DHI=TRUE } else { DHI=FALSE }

  # if working with >5 dimensions, then substitute the dimension (colnames) names with a number
  if ( d>5 ) {
    mydims<-data.frame(1:d)
    mydims$colnames<-attributes(res)$profile_dimensions
    colnames(mydims)<-c("dimension","contrast_name")
    print(kable( mydims, caption = "Profile dimensions" ))
    colnames(res$input_profile)<- paste("d",1:d,sep="")
    colnames(res$ranked_profile)<- paste("d",1:d,sep="")
    ss<-res$ranked_profile
  }

head(res$input_profile)
```

Here are some metrics about the input data profile:
```{r, metrics, echo=FALSE}
unformatted<-t(as.data.frame(res$analysis_metrics[ c(1,2,3,4,5,6,7,8,12,13 ) ])) 
formatted<-unformatted
formatted[1:4]<-as.character(round(as.numeric(unformatted[1:4]) , digits=0))
formatted[5:6]<-as.character(round(as.numeric(unformatted[5:6]) , digits=5))
colnames(formatted)="Profile metrics"
kable( formatted, caption = "Profile metrics" )
```

Here is a plot of the input profiles. Note the dynamic ranges.
```{r, scatterplot, echo=FALSE,fig.height = 6, fig.width = 6.5, message=FALSE, warning=FALSE}
if ( d==1 ) {
  par(mfrow=c(2,1))
  hist(res$input_profile[,1],breaks=50,main="Distribution of DE scores",xlab=paste("DE score for ",colnames(res$input_profile)))
  plot(res$input_profile,xlab=paste("DE score for ",colnames(res$input_profile)),
   pch="|",frame.plot=F)
  UPS=length(which(res$input_profile>0))
  DNS=length(which(res$input_profile<0))
  TOTAL=nrow(res$input_profile)
  mtext(paste(TOTAL,"genes in total,",UPS,"trending up-regulated,",DNS,"trending down-regulated"))

} else if ( d<3 ) {
  plot(res$input_profile, pch=19, col=rgb(red = 0, green = 0, blue = 0, alpha = 0.15), main="Input profiles"  )
} else {
  library("GGally")
  ggpairs_points_plot <- function(data ,mapping, ...){
    p <- ggplot(data = data, mapping = mapping) +
    geom_point(alpha=0.05) +
    geom_vline(xintercept=0,linetype="dashed") +
    geom_hline(yintercept=0,linetype="dashed")
  }
  p<-ggpairs(as.data.frame(res$input_profile), title="Scatterplot of all genes" , lower  = list(continuous = ggpairs_points_plot ))
  print( p +  theme_bw() )
}
```

Here is the contour plot of the profile including all genes.
```{r, contourplot, echo=FALSE,fig.height = 6, fig.width = 6.5,warning=FALSE, message=FALSE}
palette <- colorRampPalette(c("white", "yellow","orange" ,"red","darkred","black"))
#Contour of all the data
if (d==1) {
  message("Contour plot does not apply to unidimensional analysis.")
} else if (d==2) {
  ss<-res$ranked_profile
  xmin=min(ss[,1])
  xmax=max(ss[,1])
  ymin=min(ss[,2])
  ymax=max(ss[,2])
  ss<-res$ranked_profile
  k<-MASS:::kde2d(ss[,1],ss[,2])
  X_AXIS=paste("Rank in contrast",colnames(ss)[1])
  Y_AXIS=paste("Rank in contrast",colnames(ss)[2])
  filled.contour(k, xlim=c(xmin,xmax),ylim=c(ymin,ymax),
    color=palette ,
    plot.title={ abline(v=0,h=0,lty=2,lwd=2,col="blue")
      title( main="Rank-rank plot of all genes",xlab=X_AXIS,ylab=Y_AXIS ) } )
} else if (d>2) {
  #pairs contour plot function
  ggpairs_func <- function(data, mapping, ...){
    p <- ggplot(data = data, mapping = mapping) +
      stat_density2d(aes(fill=..density..), geom="tile", contour = FALSE) +
      geom_vline(xintercept=0,linetype="dashed") +
      geom_hline(yintercept=0,linetype="dashed") +
      scale_fill_gradientn(colours=palette(25))
    p
  }

  #pairs contour plot
  p<-ggpairs(as.data.frame(ss), title="Contour plot of all genes after ranking" , lower=list(continuous=ggpairs_func),
    diag=list(continuous=wrap("barDiag", binwidth=nrow(ss)/100)))
  print( p + theme_bw() )
  #subset contour plot
  ggpairs_contour_limit_range <- function(data ,mapping, ...){
    p <- ggplot(data = data, mapping = mapping) +
      stat_density2d(aes(fill=..density..), geom="tile", contour = FALSE) +
      geom_vline(xintercept=0,linetype="dashed") +
      geom_hline(yintercept=0,linetype="dashed") +
      scale_fill_gradientn(colours=palette(25)) +
      scale_x_continuous( limits = range(min(ss[,gsub("~","",as.character(mapping[1]))]),max(ss[,gsub("~","",as.character(mapping[1]))])) ) +
      scale_y_continuous( limits = range(min(ss[,gsub("~","",as.character(mapping[2]))]),max(ss[,gsub("~","",as.character(mapping[2]))])) )
    p
  }

  #subset points plot
  ggpairs_points_limit_range <- function(data ,mapping, ...){
    p <- ggplot(data = data, mapping = mapping) +
      geom_point(alpha=0.1) +
      geom_vline(xintercept=0,linetype="dashed") +
      geom_hline(yintercept=0,linetype="dashed") +
      scale_x_continuous( limits = range(min(ss[,gsub("~","",as.character(mapping[1]))]),max(ss[,gsub("~","",as.character(mapping[1]))])) ) +
      scale_y_continuous( limits = range(min(ss[,gsub("~","",as.character(mapping[2]))]),max(ss[,gsub("~","",as.character(mapping[2]))])) )
    p
  }
}
```

Here is the distribution of genes in the quadrants/sectors
```{r, quadrant_distribution, results="asis",echo=FALSE}
if (d==1) {

  message("A volcano plot For unidimensional analysis.")

  par(mfrow=c(1,1))
  sig<-subset(res$enrichment_result,p.adjustANOVA<=0.05)
  plot(res$enrichment_result$s.dist,-log10(res$enrichment_result$pANOVA),
   xlab="s score",ylab="-log10(p-value)",
   main="volcano plot of gene set enrichments",pch=19,cex=0.8)
  points(sig$s.dist,-log10(sig$pANOVA),pch=19,cex=0.85,col="red")
  TOTAL=nrow(res$enrichment_result)
  SIG=nrow(sig)
  UP=length(which(sig$s.dist>0))
  DN=length(which(sig$s.dist<0))
  SUBHEADER=paste(TOTAL,"gene sets in total,",UP,"upregulated and ",DN,"downregulated (FDR<=0.05)")
  mtext(SUBHEADER)

} else if ( d<3 ) {
  uu=length(which(res$input_profile[,1]>0 & res$input_profile[,2]>0))
  ud=length(which(res$input_profile[,1]>0 & res$input_profile[,2]<0))
  dd=length(which(res$input_profile[,1]<0 & res$input_profile[,2]<0))
  du=length(which(res$input_profile[,1]<0 & res$input_profile[,2]>0))
  a<-as.data.frame(c(uu,ud,dd,du))
  rownames(a)=c("top-right","bottom-right","bottom-left","top-left")
  colnames(a)="a"
  barplot(a$a,names.arg=rownames(a),main="number of genes in each quadrant")
} else {
  sig<-sign(ss)
  sector_count<-aggregate(1:nrow(sig) ~ ., sig, FUN = length)
  colnames(sector_count)[ncol(sector_count)]<-"Count"
  kable(sector_count ,caption = "Genes by sector" ,row.names=T)
}
```

## Input genesets
Here are some metrics about the gene sets used:
```{r,input_geneset_metrics1,results="asis",echo=FALSE}
ORIGINFILE=attributes(genesets)$originfile
print(paste("GMT file of genesets:",ORIGINFILE))
unformatted<-t(as.data.frame(res$analysis_metrics[ c(1,6,7,13  ) ]))
formatted<-as.data.frame(as.character( unformatted[1:4]) )
rownames(formatted)=rownames(unformatted)
colnames(formatted)="Gene sets metrics"
kable( formatted , col.names = "Gene sets metrics" ,caption = "Gene sets metrics" )
```

```{r,input_geneset_metrics2,results="asis",echo=FALSE ,fig.height = 7, fig.width = 7 ,fig.show="all"}
par(mfrow=c(3,1))
geneset_counts<-res$analysis_metrics$geneset_counts
boxplot(geneset_counts$count,horizontal=T,frame=F,main="Gene set size",xlab="number of member genes included in profile")
hist(geneset_counts$count,100,xlab="geneset size",main="Histogram of geneset size")
hist(geneset_counts$count,100,xlim=c(0,500),xlab="geneset size",main="Trimmed histogram of geneset size")
```

```{r,input_geneset_metrics3,results="asis",echo=FALSE}
if (d<2) {
nsig=length(which(res$enrichment_result$p.adjustANOVA<0.05))
} else {
nsig=length(which(res$enrichment_result$p.adjustMANOVA<0.05))
}

if (d>1) {
  a<-res$analysis_metrics[14]
  a<-as.data.frame(as.numeric(unlist(strsplit(as.character(a),','))),stringsAsFactors=F)
  rownames(a)=c("top-right","bottom-right","bottom-left","top-left")
  colnames(a)="a"
  barplot(a$a,names.arg=rownames(a),main="number of genesets FDR<0.05")
} 
if ( nsig > 0 ) {
  if ( d>1 ) {
    sig<-sign(res$enrichment_result[which(res$enrichment_result$p.adjustMANOVA<0.05),4:(4+d-1)])
    sector_count<-aggregate(1:nrow(sig) ~ ., sig, FUN = length)
    colnames(sector_count)[ncol(sector_count)]<-"Count"
    kable(sector_count ,caption = "Gene sets by sector" ,row.names=T)
    dev.off()
  }
}
```

## Enrichment scatterplot with taucharts
```{r,tauchart1d,results="asis",echo=F, fig.height = 7, fig.width = 7 ,fig.show="all", message=FALSE,eval=D1}
#1D tauchart
library("taucharts")
library(dplyr)
library(gtools)
numsets=nrow(subset(res$enrichment_result,p.adjustANOVA<0.05))
p=NULL
if (numsets==0){
  message("No significant enrichments found.")
} else {
  # volcano tauchart
  myres<-subset(res$enrichment_result,p.adjustANOVA<0.05)
  myres$confESp=NULL # NA values will screw up the tauchart
  myres$significance<--log10(myres$pANOVA)
  myres$set<-gsub("_"," ",as.character(myres$set))
  myres[,3:ncol(myres)]<-signif(myres[,3:ncol(myres)],3)
  colnames(myres)=gsub("\\.","_",colnames(myres))
  my_x=colnames(myres)[4]
  my_y=colnames(myres)[6]
  p<-tauchart(myres) %>% tau_point(my_x,my_y) %>% tau_tooltip()
  message("Significance is calculated by -log10(p-value). All points shown are FDR<0.05")
}
print("All sets with FDR<0.05. Try hovering over the points.")
p

print("Top N sets irrespective of FDR. Try hovering over the points.")
# volcano tauchart
myres<-head(res$enrichment_result,resrows)
myres$significance<--log10(myres$pANOVA)
myres$set<-gsub("_"," ",as.character(myres$set))
colnames(myres)=gsub("\\.","_",colnames(myres))
my_x=colnames(myres)[4]
my_y=colnames(myres)[7]
myres$confESp=NULL  # NA values will screw up the tauchart
tauchart(myres) %>% tau_point(my_x,my_y) %>% tau_tooltip()
message("Significance is calculated by -log10(p-value). All points shown are FDR<0.05")
```

```{r,tauchart2,results="asis",echo=F, fig.height = 7, fig.width = 7 ,fig.show="all", message=FALSE,eval=!D1}
library("taucharts")
library(dplyr)
library(gtools)

res$enrichment_result$confESp<-as.numeric(res$enrichment_result$confESp)
numsets=nrow(subset(res$enrichment_result,p.adjustMANOVA<0.05))
p=NULL
if (numsets>0){
  myres<-subset(res$enrichment_result,p.adjustMANOVA<0.05)
  # myres[,3:ncol(myres)]<-signif(myres[,3:ncol(myres)],3)
  if ( d<3 ) {
    myres<-myres[,c(1,which(names(myres) %in% "p.adjustMANOVA"),2, 4:(4+d-1) , which(names(myres) %in% "s.dist") )]
    myres$set<-gsub("_"," ",as.character(myres$set))
    colnames(myres)=gsub("\\.","_",colnames(myres))
    my_x=colnames(myres)[4]
    my_y=colnames(myres)[5]    
    tauchart(myres) %>% tau_point(my_x,my_y) %>% tau_tooltip()
  } else {
    plan<-combinations(n = d, r = 2, v = 1:d, repeats.allowed = FALSE)
    myres<-myres[,c(1, which(names(myres) %in% "p.adjustMANOVA") ,2,4:(4+d-1), which(names(myres) %in% "s.dist") )]
    p<-list()
    tauscatter<-function(i){
      my_x=plan[i,1]
      my_y=plan[i,2]
      colnames(myres)=gsub("\\.","_",colnames(myres))
      my_x=colnames(myres)[3+my_x]
      my_y=colnames(myres)[3+my_y]
      myres$set<-gsub("_"," ",as.character(myres$set))
      tauchart(myres) %>% tau_point(my_x,my_y) %>% tau_tooltip() 
    }
    p<-lapply( 1:nrow(plan) , tauscatter)
  }
} else {
  message("No significant enrichments found.")
}
    #output a bunch of taucharts doesnt work as expected
    #p

    #this doesn't work either
    #while ( length(p)>0 ) {
    #  pp<-p[[1]] ; p[[1]]=NULL ; pp                
    #}

  #need to use this ugly hack to print plots
  #run the following 120 times which is 15 nth triangular number
  #pp<-p[[1]] ; p[[1]]=NULL ; pp

  # try this suggestion 
  # do.call(subplot, lapply(colnames(mydf)[2:3], function(column_name) {
  #   plot_ly(x = mydf$x, y = mydf[,column_name])
  # }))

if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }

if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }

if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }

if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }

if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }

if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }

if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }

if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }

if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }

if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }

if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }

if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
}

print("Top N sets irrespective of FDR. Try hovering over the points.")
message(paste("Prioritisation done by" , attributes(res$enrichment_result)$priority ))
resrows=length(res$detailed_sets)
myres<-head(res$enrichment_result,resrows)
myres[,3:ncol(myres)]<-signif(myres[,3:ncol(myres)],3)

if ( d<3 ) {
  myres[,3:ncol(myres)]<-signif(myres[,3:ncol(myres)],3)
  myres<-myres[,c(1,which(names(myres) %in% "p.adjustMANOVA"),2, 4:(4+d-1) , which(names(myres) %in% "s.dist") )]
  myres$set<-gsub("_"," ",as.character(myres$set))
  colnames(myres)=gsub("\\.","_",colnames(myres))
  my_x=colnames(myres)[4]
  my_y=colnames(myres)[5]
  tauchart(myres) %>% tau_point(my_x,my_y) %>% tau_tooltip()
} else {
  plan<-combinations(n = d, r = 2, v = 1:d, repeats.allowed = FALSE)
  myres<-myres[,c(1, which(names(myres) %in% "p.adjustMANOVA") ,2,4:(4+d-1), which(names(myres) %in% "s.dist") )]
  p<-list()
  tauscatter<-function(i){
    my_x=plan[i,1]
    my_y=plan[i,2]
    colnames(myres)=gsub("\\.","_",colnames(myres))
    my_x=colnames(myres)[3+my_x]
    my_y=colnames(myres)[3+my_y]
    myres$set<-gsub("_"," ",as.character(myres$set))
    tauchart(myres) %>% tau_point(my_x,my_y) %>% tau_tooltip()
  }
  p<-lapply( 1:nrow(plan) , tauscatter)
}

if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }

if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }

if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }

if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }

if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }

if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }

if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }

if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }

if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }

if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }

if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }

if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
if ( length(p)>0 ) { pp<-p[[1]] ; p[[1]]=NULL ; pp }
}
```

```{r,heatmap,results="asis",echo=F, fig.height = 7, fig.width = 7 ,fig.show="all", message=FALSE,eval=!D1}
asis_output("## A heatmap of S values for top results\\n")
hmapx<-head( res$enrichment_result[,4:(4+d-1)] ,resrows)
rownames(hmapx)<-head(res$enrichment_result$set,resrows)
colnames(hmapx)<-gsub("^s.","",colnames(hmapx))
my_palette <- colorRampPalette(c("blue", "white", "red"))(n = 25)
heatmap.2(as.matrix(hmapx),scale="none",margin=c(10, 25),cexRow=0.8,trace="none",cexCol=0.8,col=my_palette)
```

```{r,effectsize,results="asis",echo=F, fig.height = 7, fig.width = 7 ,fig.show="all", message=FALSE,eval=!D1}
asis_output("## A plot of effect size versus significance\\n")
library(gplots)
myres<-res$enrichment_result
myres[,3:ncol(myres)]<-signif(myres[,3:ncol(myres)],3)
myres$significance<--log2(myres$p.adjustMANOVA)
myres$set<-gsub("_"," ",as.character(myres$set))
colnames(myres)=gsub("\\.","_",colnames(myres))
my_x=colnames(myres)[ncol(myres)-2]
my_y=colnames(myres)[ncol(myres)]
tauchart(myres) %>% tau_point(my_x,my_y) %>% tau_tooltip()
print("Significance is the -log2(p.adjustMANOVA) and effect size is the s.dist which is the hypotenuse of the s scores.")
```

## Results table
```{r,results_table,results="asis",echo=FALSE}
if ( d==1 ) {
  resrows=length(res$detailed_sets)
  myres<-head(res$enrichment_result,resrows)
  myres$confESp=NULL
  myres[,c(3:ncol(myres))]<-signif(myres[,c(3:ncol(myres))],3)
} else if (d>1) {
  resrows=length(res$detailed_sets)
  myres<-head(res$enrichment_result,resrows)
  myres<-myres[,c(1:3, which(names(myres) %in% "p.adjustMANOVA"), which(names(myres) %in% "s.dist"), which(names(myres) %in% "confESp") ,4:((d*2)+3))]
  myres[,3:ncol(myres)]<-signif(myres[,3:ncol(myres)],3)
}
formatted<-myres
formatted[,1]<-gsub("_"," ",myres[,1])
kable( formatted , col.names = colnames(formatted) , row.names=F, caption = "Top gene sets" ,digits=100)
cat("<hr><br>")
```

## Results (complete table)
```{r,results_table_complete,results="asis",echo=FALSE}
myres<-res$enrichment_result

if ( d==1 ) {
  myres<-res$enrichment_result
  myres$confESp=NULL
  myres[,c(3:ncol(myres))]<-signif(myres[,c(3:ncol(myres))],3)
  formatted<-myres
  formatted[,1]<-gsub("_"," ",myres[,1])
} else if (d>1) {
  myres<-myres[,c(1:3, which(names(myres) %in% "p.adjustMANOVA"), which(names(myres) %in% "s.dist"), which(names(myres) %in% "confESp") ,4:((d*2)+3))]
  myres[,3:ncol(myres)]<-signif(myres[,3:ncol(myres)],3)
  formatted<-myres
  formatted[,1]<-gsub("_"," ",myres[,1])
  formatted[,3:ncol(myres)]<-format(myres[,3:ncol(myres)],digits=3)
}
HEADER=paste("<br><details><summary><b>","Click HERE to show results for all gene sets","</b></summary><br><p>",sep=" " )
cat(HEADER)
kable(formatted, col.names=colnames(formatted),row.names=F,caption = "Complete results",digits=100)
cat("<br></p></details>")
cat("<hr><br>")
```

```{r,detailed_geneset_reports1d,results="asis",echo=FALSE,fig.height=6, fig.width=6 ,out.width = '80%',comment=NA, message=FALSE,eval=D1}
cat("## Detailed Gene set reports\\n")
cat('<br>')
cat('\n')
library("beeswarm")
for ( i in 1:resrows ) {
  mydat = NULL
  mydat<-t(myres[i,])
  cat(paste("<b>",as.character(myres[i,1]) ,"</b><br>"))
  print(kable(mydat,format='markdown',caption=as.character(t(myres[i,1]))) )
  cat('\n')
  cat("<br>")

  # plots
  sss<-res$detailed_sets[[i]]
  set<-names(res$detailed_sets[i])
  size<-length(sss)

  par(mfrow=c(3,1))
  beeswarm(sss,vertical = F,cex=0.75,xlim=c(min(ss),
   max(ss)),col="darkgray",pch=19,main=set,cex.main=1.5,
   xlab=paste("ranked DE score in:",colnames(ss)))
   mtext("beeswarm plot",cex=0.8)
  hist(sss,xlim=c(min(ss),max(ss)),breaks=15,col="darkgray",main=NULL,
   border="black",xlab=paste("ranked DE score in:",colnames(ss)))
   mtext("histogram",cex=0.8)
  plot(sss,rep(1,length(sss)),type="n",xlim=c(min(ss),max(ss)),
   frame=F,axes=F,ylab="",
   xlab=paste("ranked DE score in:",colnames(ss)))
  rug(sss, ticksize = 0.9)
  axis(1)
  mtext("rugplot",cex=0.8)

  cat("<br>")
  # top gene list
  setSign=sign(res$enrichment_result[i,4])

  if (setSign==-1) {
    tops<-as.data.frame(res$detailed_sets[[i]][order(res$detailed_sets[[i]])])
  } else if (setSign==1) {
    tops<-as.data.frame(res$detailed_sets[[i]][order(-res$detailed_sets[[i]])])
  }
cat("Top enriched genes\\n")
tops$GeneID<-rownames(tops)
tops<-tops[,c(2,1)]
colnames(tops)=c("GeneID","Gene Rank")
cat("<br>")
print(kable(head(tops,n=20L),col.names=colnames(tops),format="markdown",row.names=F,caption="Top 20 genes",digits=100))
cat('\n')

HEADER=paste("<br><details><summary><b>","Click HERE to show all gene set members","</b></summary><br><p>",sep=" " )
cat(HEADER)
print(kable(tops,col.names=colnames(tops),format="markdown",row.names=F,caption="All member genes",digits=100))
cat('\n')
cat("<br></p></details>")
cat("<br><hr>")
}
```



```{r,detailed_geneset_reports2d,results="asis",echo=FALSE,fig.height=5, fig.width=6 ,out.width = '100%',comment=NA, message=FALSE,eval=!D1}
asis_output("## Detailed Gene set reports\\n")
library("reshape2")
library("ggplot2")
library("GGally")
resrows=length(res$detailed_sets)
ss<-res$ranked_profile
myres<-head(res$enrichment_result,resrows)


myres<-myres[,c(1:3, which(names(myres) %in% "p.adjustMANOVA"), which(names(myres) %in% "s.dist"), which(names(myres) %in% "confESp") ,4:((d*2)+3))]
myres[,3:ncol(myres)]<-signif(myres[,3:ncol(myres)],3)
palette <- colorRampPalette(c("white", "yellow","orange" ,"red","darkred","black"))
ss_long<-melt(ss)

twodimplot<-function(i) {
  ll<-res$enrichment_result[i,]
  size<-ll$setSize
  sss<-res$detailed_sets[[i]]
  k<-MASS:::kde2d(sss[,1],sss[,2])
   filled.contour( k, color = palette, xlim=c(xmin,xmax),ylim=c(ymin,ymax),
   plot.title={ abline(v=0,h=0,lty=2,lwd=2,col="blue")
   title( main=ll$set , xlab=X_AXIS,ylab=Y_AXIS )})

  plot(sss, pch=19, col=rgb(red = 0, green = 0, blue = 0, alpha = 0.2),
   main=ll$set ,
   xlim=c(xmin,xmax),ylim=c(ymin,ymax),
   xlab=X_AXIS,ylab=Y_AXIS)

  abline(v=0,h=0,lty=2,lwd=2,col="blue")

  sss_long<-melt(sss)

  p<-ggplot(ss_long,aes(Var2,value)) +
   geom_violin(data=ss_long,fill = "grey", colour = "grey") +
   geom_boxplot(data=ss_long,width=0.9,fill="grey",outlier.shape = NA) +
   geom_violin(data=sss_long,fill = "black", colour = "black") +
   geom_boxplot(data=sss_long,width=0.1,outlier.shape = NA) +
   labs(y = "Position in rank",title = ll[,1] )

  print( p + theme_bw() +
   theme( axis.text=element_text(size=14),
   axis.title=element_text(size=15),
   plot.title = element_text(size = 20)))
}

ndplot<-function(i) {
  ll<-res$enrichment_result[i,]
  size<-ll$setSize
  ss<-res$ranked_profile
  sss<-res$detailed_sets[[i]]

  if ( d>5 ) {
    colnames(sss)<- paste("d",1:d,sep="")
  }

  #subset contour plot
  ggpairs_contour_limit_range <- function(data ,mapping, ...){
    p <- ggplot(data = data, mapping = mapping) +
     stat_density2d(aes(fill=..density..), geom="tile", contour = FALSE) +
     geom_vline(xintercept=0,linetype="dashed") +
     geom_hline(yintercept=0,linetype="dashed") +
     scale_fill_gradientn(colours=palette(25)) +
     scale_x_continuous( limits = range(min(ss[,gsub("~","",as.character(mapping[1]))]),max(ss[,gsub("~","",as.character(mapping[1]))])) ) +
     scale_y_continuous( limits = range(min(ss[,gsub("~","",as.character(mapping[2]))]),max(ss[,gsub("~","",as.character(mapping[2]))])) )
    p
  }

  #subset points plot
  ggpairs_points_limit_range <- function(data ,mapping, ...){
    p <- ggplot(data = data, mapping = mapping) +
     geom_point(alpha=0.1) +
     geom_vline(xintercept=0,linetype="dashed") +
     geom_hline(yintercept=0,linetype="dashed") +
     scale_x_continuous( limits = range(min(ss[,gsub("~","",as.character(mapping[1]))]),max(ss[,gsub("~","",as.character(mapping[1]))])) ) +
     scale_y_continuous( limits = range(min(ss[,gsub("~","",as.character(mapping[2]))]),max(ss[,gsub("~","",as.character(mapping[2]))])) )
    p
  }

  p<-ggpairs(as.data.frame(sss), title=ll[,1], lower=list(continuous=ggpairs_contour_limit_range),
   diag=list(continuous=wrap("barDiag", binwidth=nrow(ss)/10)) )
  print( p + theme_bw() )
  p<-ggpairs(as.data.frame(sss), title=ll[,1], lower= list(continuous = ggpairs_points_limit_range ),
   diag=list(continuous=wrap("barDiag", binwidth=nrow(ss)/10)))
  print( p + theme_bw() )

  sss_long<-melt(sss)

  p<-ggplot(ss_long,aes(Var2,value)) +
   geom_violin(data=ss_long,fill = "grey", colour = "grey") +
   geom_boxplot(data=ss_long,width=0.9,fill="grey",outlier.shape = NA) +
   geom_violin(data=sss_long,fill = "black", colour = "black") +
   geom_boxplot(data=sss_long,width=0.1,outlier.shape = NA) +
   labs(y = "Position in rank",title = ll[,1] )

  print( p + theme_bw() +
   theme( axis.text=element_text(size=14),
   axis.title=element_text(size=15),
   plot.title = element_text(size = 20)))

}

topgenelist<-function(i) {
  ss<-res$ranked_profile
  sss<-res$detailed_sets[[i]]
  if ( d>5 ) {
    colnames(sss)<- paste("d",1:d,sep="")
  }

  tl=bl=tr=br=0
  myres<-head(res$enrichment_result,resrows)
  myres<-myres[,c(1:3, which(names(myres) %in% "p.adjustMANOVA"), which(names(myres) %in% "s.dist"), which(names(myres) %in% "confESp") ,4:((d*2)+3))]
  myres[,3:ncol(myres)]<-signif(myres[,3:ncol(myres)],3)

  #select 2 strongest columns to highlight genes
  if( d>2 ) {
    cols<-(ncol(myres)-(2*d)+1):(ncol(myres)-(2*d)+d)
    COLS<-order(-abs(myres[i,cols]))[1:2]
    sss<-sss[,COLS]
    cols<-cols[COLS]
    mysx=signif(myres[i,cols[1]],3)
    mysy=signif(myres[i,cols[2]],3)
  } else {
    cols=c(ncol(myres)-3,ncol(myres)-2)
    mysx=signif(myres[i,cols[1]],3)
    mysy=signif(myres[i,cols[2]],3)
  }

  if ( mysy>0 ) { tl=tl+1 ; tr=tr+1  } else { bl=bl+1 ; br=br+1 }
  if ( mysx>0 ) { tr=tr+1 ; br=br+1  } else { tl=tl+1 ; bl=bl+1 }
  if (bl==2) {
    myquad<-sss[which(sss[,1]<0 & sss[,2]<0),]
    topgenes<-myquad[order(-rank(myquad[,1]*myquad[,2])),]
  }
  if (tr==2) {
    myquad<-sss[which(sss[,1]>0 & sss[,2]>0),]
    topgenes<-myquad[order(-rank(myquad[,1]*myquad[,2])),]
  }
  if (br==2) {
    myquad<-sss[which(sss[,1]>0 & sss[,2]<0),]
    topgenes<-myquad[order(rank(myquad[,1]*myquad[,2])),]
  }
  if (tl==2) {
    myquad<-sss[which(sss[,1]<0 & sss[,2]>0),]
    topgenes<-myquad[order(rank(myquad[,1]*myquad[,2])),]
  }
  cat("<br>")
  topgenes<-as.data.frame(topgenes)
  topgenes$Gene<-as.character(rownames(topgenes))
  topgenes<-topgenes[,c(3,1,2)]
  print(kable(head(topgenes,n=20L),col.names=colnames(topgenes),format="markdown",row.names=F,caption="Top 20 genes",digits=100))
  cat('\n')
  HEADER=paste("<br><details><summary><b>","Click HERE to show all gene set members","</b></summary><br><p>",sep=" " )
  cat(HEADER)
  sss<-res$detailed_sets[[i]]
  print(kable(sss,format="markdown",row.names=T,caption="All member genes",digits=100))
  cat('\n')
  cat("<br></p></details>")
  cat("<br><hr>")
}

ss_long<-melt(ss)

#functions end here
if ( d<3 ) {
  xmin=min(ss[,1])
  xmax=max(ss[,1])
  ymin=min(ss[,2])
  ymax=max(ss[,2])
  for ( i in 1:resrows ) {
    mydat = NULL                                                    
    mydat$metrics<-names(myres[2:ncol(myres)])
    mydat$values<-unname(t(myres[i,2:ncol(myres)]))
    mydat<-as.data.frame(cbind(mydat$metrics,mydat$values))
    colnames(mydat)<-c("metric","value")
    cat(paste("<b>",as.character(myres[i,1]) ,"</b><br>"))
    print(kable(mydat,format='markdown',caption=as.character(t(myres[i,1]))) )
    cat('\n')
    cat("<br>")
    twodimplot(i)
    cat("<br>")
    topgenelist(i)
    cat("<hr><br>")
  }
} else {
  for ( i in 1:resrows ) {
    mydat = NULL
    mydat$metrics<-names(myres[2:ncol(myres)])
    mydat$values<-unname(t(myres[i,2:ncol(myres)]))
    mydat<-as.data.frame(cbind(mydat$metrics,mydat$values))
    colnames(mydat)<-c("metric","value")
    cat(paste("<b>",as.character(myres[i,1]) ,"</b><br>"))
    print(kable(mydat,format='markdown',digits=5,caption=as.character(t(myres[i,1]))))
    cat('\n')
    cat("<br>")
    ndplot(i)
    cat("<br>")
    topgenelist(i)
    cat("<hr><br>")
  }
}
```

Here is the session info with all the versions of packages used.

```{r session_info, include=TRUE, echo=TRUE, results='markup'}
devtools::session_info()
```

END of report
