---
title: "Mitch Workflow"
author: "Mark Ziemann & Antony Kaspi"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Mitch Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

## Background

Mitch performs multidimensional gene set enrichment analysis. The concept behind this dates to work by Cox and 
Mann (https://doi.org/10.1186/1471-2105-13-S16-S12). This implementation is suited to R based workflows of 
multi-omics datasets. This software was developed by Antony Kaspi and Mark Ziemann. Learn
more about Mitch at the website: https://github.com/markziemann/Mitch

```{r,lib,echo=FALSE}
library("mitch")
```

### Workflow overview
Importing gene sets
Mitch has a function to import GMT files to R lists (adapted from Yu et al, 2012 in the clusterProfiler package). For example:

```{r,gsets,echo=FALSE}
download.file("https://reactome.org/download/current/ReactomePathways.gmt.zip",destfile="ReactomePathways.gmt.zip")
unzip("ReactomePathways.gmt.zip")
genesets<-gmt_import("ReactomePathways.gmt")
```

### Importing profiling data
Mitch accepts pre-ranked data supplied by the user, but also has a function called mitch_import for importing tables generated by Limma, edgeR, DESeq2, ABSSeq, Sleuth, Seurat and Muscat. By default, only the genes that are detected in all contrasts are included, but this behaviour can be modified. The below example imports two edgeR tables called "dge1" and "dge2". Where gene identifiers are coded in the row names.

```{r,import1,echo=FALSE}
x<-list("dge1"=dge1,"dge2"=dge2)
y<-mitch_import(x,"edger")
```

Mitch can do unidimensional analysis if you provide it a single profile as a dataframe.

```{r,import2,echo=FALSE}
y<-mitch_import(df,DEtype="edger")
```

If the gene identifiers are not given in the rownames, then the column can be specified with the geneIDcol parameter like this:

```{r,import3,echo=FALSE}
y<-mitch_import(df,DEtype="edger",geneIDcol="MyGeneIDs")
```

By default, differential gene activity is scored using the directional nominal p-value.

S = -log10(p-value) * sgn(logFC)

If this is not desired, then users can perform their own custom scoring procedure.

There are many cases where the gene IDs don't match the gene sets. To overcome this, mitch_import also accepts a two-column table that relates gene identifiers in the profiling data to those in the gene sets.

`?mitch_import` provides more instructions on using this feature.

### Calculating enrichment
The mitch_calc function performs multivariate enrichment analysis of the supplied gene sets in the scored profiling data. At its simpest form mitch_calc function accepts the scored data as the first argument and the genesets as the second argument. Users can prioritise enrichments based on small adjusted p-values, or by the observed effect size (magnitude of the enrichment score).

```{r,calc1,echo=FALSE}
res<-mitch_calc(y,genesets,priority="significance")
res<-mitch_calc(y,genesets,priority="effect")
```
You can peek at the top results with head like this:

`head(res$enrichment_result)`

By default, mitch_calc uses mclapply to speed up calculations on all but one available CPU threads. This behaviour can be modified by setting the cores to a desred number.

```{r,calc2,echo=FALSE}
res<-mitch_calc(y,genesets,priority="significance",cores=4)
```
By default, gene sets with fewer than 10 members present in the profiling data are discarded. This threshold can be modified using the minsetsize option. There is no upper limit of gene set size.

```{r,calc3,echo=FALSE}
res<-mitch_calc(y,genesets,priority="significance",minsetsize=20)
```
By default, in downstream visualisation steps, charts are made from the top 50 gene sets, but this can be modified using the resrows option.

```{r,calc4,echo=FALSE}
res<-mitch_calc(y,genesets,priority="significance",resrows=100)
```
Generate a HTML report
Can be done like this:

```{r,report,echo=FALSE}
mitch_report(res,"myreport.html")
```
If you receive an error here, ensure that Taucharts is installed properly. Take a look at an example report.

Generate high resolution plots
In case you want the charts in PDF format, these can be generated as such:

```{r,plots,echo=FALSE}
mitch_plots(res,outfile="mycharts.pdf")
```
